# Reproducible Research: Peer Assessment 1

## Loading libraries
```{r, echo=TRUE}
library(knitr)
library(lattice)
```

## Loading and preprocessing the data
```{r}
data <- read.csv("activity.csv")
#data_total_steps <- 
#  data.frame(total.steps=rowsum(data$steps,data$date,na.rm=T))
#data_total_steps$date <- rownames(data_total_steps)
#data.2 <- merge(data,data_total_steps,by="date")
step_sum <- function(data){
  sum_steps <- c(); current_date <- ""
  steps <- data[,"steps"]; date <- data[,"date"]
  for (i in 1:nrow(data)){
    if (current_date != date[i]){
      current_date <- date[i]
      sum_steps[i] <- steps[i]
    } else {
      sum_steps[i] <- sum_steps[i-1] + steps[i]
    }
  }
  return(sum_steps)
}
data$sum_step <- step_sum(data)
data$ave_step <- data$sum_step / (data$interval + 5)
##data.na_omit <- data[!is.na(data$steps),]
```


## What is mean total number of steps taken per day?
```{r,echo=TRUE}
hist(rowsum(data$steps,data$date,na.rm=T), 
     main="Total number of steps taken per day",
     xlab="Total steps per day")
step_total_mean = mean(rowsum(data$steps,data$date,na.rm=T))
step_total_median = median(rowsum(data$steps,data$date,na.rm=T))
print(sprintf("mean: %f", step_total_mean))
print(sprintf("median: %d",step_total_median))
```
The mean and median total number of steps taken per day are 
`r step_total_mean` and `r step_total_median` respectively.

## What is the average daily activity pattern?
```{r}
mean_omitna <- function(data){mean(data,na.rm=T)}
avg_step <- tapply(data$steps, data$interval, mean_omitna)
plot(as.integer(names(avg_step)), avg_step,type="l")
avg_step[avg_step==max(avg_step)]


#plot(x=data$interval, y=data$ave_step,
#     type="l")
#data[data$ave_step==max(data$ave_step,na.rm=T)  
#       & !is.na(data$sum_step),]
```

## Imputing missing values
```{r,echo=TRUE}
n_na = sum(is.na(data$steps) )
print(n_na)

# NA filling by corresponding interval mean
mean_interval <- tapply(data$steps, data$interval, mean_naomit <- function(x){ mean(x,na.rm=T)})
data.tmp <- data.frame(interval=as.integer(names(mean_interval)), mean_steps=mean_interval)
data.fill_na <- merge(data, data.tmp, by="interval", sort=T)
data.fill_na <- data.fill_na[order(data.fill_na$date, data.fill_na$interval), ]
data.fill_na[is.na(data.fill_na$steps),"steps"] <- 
  data.fill_na[is.na(data.fill_na$steps), "mean_steps"]
data.fill_na <- data.fill_na[,c("steps","date","interval")]
rownames(data.fill_na) <- seq(1,nrow(data.fill_na))

# create histogram & calculate mean/median
hist(rowsum(data.fill_na$steps,data.fill_na$date,na.rm=T), 
     main="Total number of steps taken per day",
     xlab="Total steps per day")
step_total_mean = mean(rowsum(data.fill_na$steps,data.fill_na$date,na.rm=T))
step_total_median = median(rowsum(data.fill_na$steps,data.fill_na$date,na.rm=T))
print(sprintf("mean: %f", step_total_mean))
print(sprintf("median: %f",step_total_median))
```
The total number of missing value in the dataset is `r n_na`.

## Are there differences in activity patterns between weekdays and weekends?
```{r,echo=TRUE}
# In this section, weekdays depend on computer environment.
# My computer is under Japanese env., so the weekdays() returns Japanese.
type_weekdays <- function(weekdays){
  tmp <- function(weekday){
    x <- c("weekday", "weekend")
    fc <- factor(x, levels=c("weekday","weekend"))
    if (weekday=="0"||weekday=="6") {
        return(fc[2])
      } else {
        return(fc[1])
      }
  }
  sapply(weekdays,tmp)
}
data.fill_na$weekdays <- as.POSIXlt(as.Date(data.fill_na$date,"%Y-%m-%d"))$wday
data.fill_na$day_type <- type_weekdays(data.fill_na$weekdays)
data.weekday <- subset(data.fill_na, 
                       data.fill_na$day_type=="weekday")
data.weekend <- subset(data.fill_na,
                       data.fill_na$day_type=="weekend")
avg_step_weekday <- tapply(data.weekday$steps,
                           data.weekday$interval,
                           mean_omitna)
avg_step_weekend <- tapply(data.weekend$steps,
                           data.weekend$interval,
                           mean_omitna)
avg_step_data <- data.frame(
  steps=c(avg_step_weekday,avg_step_weekend),
  interval=c(names(avg_step_weekday),names(avg_step_weekend)),
  day_type=c(rep(type_weekdays(1),length(avg_step_weekday)),
             rep(type_weekdays(0),length(avg_step_weekend)))
  )
par(mfrow=c(2,1))
plot(as.integer(names(avg_step_weekday)),
     avg_step_weekday,type="l", main="weekday",
     xlab="interval", ylab="mean of number of steps")
plot(as.integer(names(avg_step_weekend)),
     avg_step_weekend,type="l", main="weekend",
     xlab="interval",  ylab="mean of number of steps")

#xyplot(avg_step_data$steps ~ avg_step_data$interval |
#         avg_step_data$day_type, layout=c(1,2),
#       type="l")


#data.fill_na$sum_step <- step_sum(data.fill_na)
#data.fill_na$ave_step <- data.fill_na$sum_step / (data.fill_na$interval + 5)
#xyplot(data.fill_na$ave_step ~ data.fill_na$interval | 
#         data.fill_na$day_type, layout=c(1,2))
```

