# Reproducible Research: Peer Assessment 1

## Loading libraries
```{r, echo=TRUE}
library(knitr)
library(lattice)
```

## Loading and preprocessing the data
```{r}
data <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
```{r,echo=TRUE}
hist(rowsum(data$steps,data$date,na.rm=T), 
     main="Total number of steps taken per day",
     xlab="Total steps per day")
step_total_mean = mean(rowsum(data$steps,data$date,na.rm=T))
step_total_median = median(rowsum(data$steps,data$date,na.rm=T))
print(sprintf("mean: %f", step_total_mean))
print(sprintf("median: %d",step_total_median))
```
The mean and median total number of steps taken per day are 
`r step_total_mean` and `r step_total_median` respectively.

## What is the average daily activity pattern?
```{r}
mean_omitna <- function(data){mean(data,na.rm=T)}
avg_step <- tapply(data$steps, data$interval, mean_omitna)
plot(as.integer(names(avg_step)), avg_step,type="l")
interval_max <- names(avg_step[avg_step==max(avg_step)])
avg_step[avg_step==max(avg_step)]
```
`r interval_max` has the maximum number of steps on average across all the days in the dataset.

## Imputing missing values
```{r,echo=TRUE}
n_na = sum(is.na(data$steps) )
print(n_na)

# NA filling by corresponding interval mean
mean_interval <- tapply(data$steps, data$interval, mean_naomit <- function(x){ mean(x,na.rm=T)})
data.tmp <- data.frame(interval=as.integer(names(mean_interval)), mean_steps=mean_interval)
data.fill_na <- merge(data, data.tmp, by="interval", sort=T)
data.fill_na <- data.fill_na[order(data.fill_na$date, data.fill_na$interval), ]
data.fill_na[is.na(data.fill_na$steps),"steps"] <- 
  data.fill_na[is.na(data.fill_na$steps), "mean_steps"]
data.fill_na <- data.fill_na[,c("steps","date","interval")]
rownames(data.fill_na) <- seq(1,nrow(data.fill_na))

# create histogram & calculate mean/median
hist(rowsum(data.fill_na$steps,data.fill_na$date,na.rm=T), 
     main="Total number of steps taken per day",
     xlab="Total steps per day")
step_total_mean = mean(rowsum(data.fill_na$steps,data.fill_na$date,na.rm=T))
step_total_median = median(rowsum(data.fill_na$steps,data.fill_na$date,na.rm=T))
print(sprintf("mean: %f", step_total_mean))
print(sprintf("median: %f",step_total_median))
```
The total number of missing value in the dataset is `r n_na`.
In mean calculation, the NA-filling dataset slightly impacts 
the result compared to the row dataset.
However, the median is not affected so much.

## Are there differences in activity patterns between weekdays and weekends?
```{r,echo=TRUE}
type_weekdays <- function(weekdays){
  tmp <- function(weekday){
    x <- c("weekday", "weekend")
    fc <- factor(x, levels=c("weekday","weekend"))
    if (weekday=="0"||weekday=="6") {
        return(fc[2])
      } else {
        return(fc[1])
      }
  }
  sapply(weekdays,tmp)
}
data.fill_na$weekdays <- as.POSIXlt(as.Date(data.fill_na$date,"%Y-%m-%d"))$wday
data.fill_na$day_type <- type_weekdays(data.fill_na$weekdays)
data.weekday <- subset(data.fill_na, 
                       data.fill_na$day_type=="weekday")
data.weekend <- subset(data.fill_na,
                       data.fill_na$day_type=="weekend")
avg_step_weekday <- tapply(data.weekday$steps,
                           data.weekday$interval,
                           mean_omitna)
avg_step_weekend <- tapply(data.weekend$steps,
                           data.weekend$interval,
                           mean_omitna)
avg_step_data <- data.frame(
  steps=c(avg_step_weekday,avg_step_weekend),
  interval=c(names(avg_step_weekday),names(avg_step_weekend)),
  day_type=c(rep(type_weekdays(1),length(avg_step_weekday)),
             rep(type_weekdays(0),length(avg_step_weekend)))
  )
par(mfrow=c(2,1))
plot(as.integer(names(avg_step_weekday)),
     avg_step_weekday,type="l", main="weekday",
     xlab="interval", ylab="mean of number of steps")
plot(as.integer(names(avg_step_weekend)),
     avg_step_weekend,type="l", main="weekend",
     xlab="interval",  ylab="mean of number of steps")
```
